---
title: "Week 4"
author: "Lauren Dirix"
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: 72
---

# Preparation before class

## Adjusting my research questions

Based on the discussions last week, I think I will keep my research
questions the same. Thus my research questions are: 1. Are students who
are connected to each other more similar with regards to civic
engagement, than students who are not connected to each other? 2. Do
students become more similar to each other with regards to civic
engagement, when they are connected in a network?

<br>

## What data do I need?

Here I will further describe the data that I need to answer my research
questions. The items I chose last week, still seem like the best choice
to me for measuring the concepts, I will just go into them a bit more
deeply this week, answering what level data the items measure.

**Civic engagement**

For civic engagement, the questionnaire measures both beliefs and
behaviour. I am most interested in behaviour, so I will be using that
scale from now on. This scale consists of six items, each measuring a
different civic behaviour. This is a measure on an individual, actor
level. This is a dynamic measure, it can change over time, which is why
it is interesting whether it changes because of the ties people have.

**Strong/ weak ties**

The measure for both strong and weak ties, are inherently dyadic data,
as you cannot have a tie on your own. The questions asked in the MRCS
dataset ask for directed ties, as the question is about whether actor A
friends (or study partners) with actor B, however, it is possible that
actor B does not nominate actor A, leading to a directed tie.

**Covariates**

Last week I did not look into covariates that should be considered, but
the dataset contains all the 'standard' covariates, that would be
important to control for: - Gender (RXge) - Ethnicity (RXreth) - What
year of college (Rfrstyr) These are all variables measured at the
individual level, with the first two being stable characteristics, and
the last one being dynamic.

------------------------------------------------------------------------

# In class

## SNASS tutorial

```{r}
require("igraph")
g <- make_graph("Zachary")
plot(g)
```

Visually, this isn't the same graph as in the tutorial, but the ties are
the same, and the centre is relatively the same. Apparently, the figure
is made to put the 'popular' people in the centre. Based on this, we can
say something about network size, the fact that the ties are undirected
(but is that igraph, or the data that is undirected --\> check matrix).

```{r}
gmat <- as_adjacency_matrix(g, type = "both", sparse = FALSE)
# gmat
isSymmetric(gmat)
```

Based on the matrix, there are undirected ties, because it is symmetric.
You can also check this with isSymemetric, if that is true, there is an
undirected network.

### Descriptive statistics

Transitivity

```{r}
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero"))
```

There is not one actor who is specifically very unique in their amount
of transitive ties, thus doesnt seem to be that interesting of a
statistic

Betweenness

```{r}
igraph::betweenness(g, directed = FALSE)
```

This shows that number 1 and number 34 have very high number, in the
plot you do not see that (big) difference between 33 and 34, thus we
need to recreate the plot. It seems like 1, 3, 32, 33, and 34 are
interesting, as they have large numbers.

How can we plot that?

- Colour code --\> upalate to show the brighter the colour the more
  betweenness

- Size increase --\> Size nodes dependent on betweenness

- Layout broader --\> No overlapping nodes

### Dyad census

```{r}
igraph::dyad_census(g)
```

Because it is an undirected graph, the asymmetry is always 0

### Triad census

```{r}
igraph::triad_census(g)
# I will use sna because it shows the names of the triads as well.
sna::triad.census(gmat)
unloadNamespace("sna")  #I will detach this package again, otherwise it will interfere with all kind of functions from igraph, and my students will hate me for that.
```

This shows per sort of triad, how many times it is existent in the
network

How many ways there are to close a triad, is calculated through an actor
perspective --\> If there is a triad, there are three ways to be a
triad, from each actor one

Transitivity: How many triads are closed divided by the amount of triads
that are closed or can be closed OR how many 300 networks are there,
divided by 300 + 201 networks (because only 201 can be closed) This same
goes for reciprocity (on dyad level), which cannot be done on this
network, because it has undirected ties

Local transitivity: The transitivity of one ego --\> shows you how many
transitive ties one ego has Can tell you something about how two people
with the same amount of ties, differ in their network structure

```{r}
# Caluclating transitivity in an r-function
igraph::transitivity(g, type = "global")
# Or
sna::gtrans(gmat)

# Calculating transitivity manually
triad_g <- data.frame(sna::triad.census(gmat))

transitivity_g <- (3 * triad_g$X300)/(triad_g$X201 + 3 * triad_g$X300)
transitivity_g
```

### Network visualisation

```{r}
# changing V
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 10  #after some trial and error
plot(g, mode = "undirected")
```

Meaning of the different parts of the codes:

- V = virtuses (the nodes)

- E = edges

- \*60 = betweenness score is between 0 and 1, needs to be more heavily
  taken into account

- \+10 = if you don't do that, nodes with a betweenness of 0 will also
  be visualised with 0, which would be undesirable

```{r}
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)
```
```{r}
l  #let us take a look at the coordinates
l[1, 1] <- 4
l[34, 1] <- -3.5
plot(g, layout = l)
```
This is the same network, but the second one looks/ feels a lot different (seems to suggest way more polarisation, where 1 and 34 are polar opposites)
For the second one, we manipulated the graph manually, if you do this, be honest about it!!! BUT this is the same as the algorithm does, which is something that is generally accepted, which is kind of odd. Solution: Maybe mention what algorithm you used (and why?) --> graphs aren't neutral

<br>

# Trying on my own
Load in data
```{r}
packages = c("stringdist", "stringi", "tidyverse")
fpackage.check(packages)

load("./data/IndividualCharacteristics.RData")
load("./data/NetworkCharacteristics.RData")
```

<br>

## Variables of interest
### Network matrix (weak ties)
Create a matrix, that can be used within RSiena. I set all values 9 to NA, as these respondents have not filled out the survey

```{r}
# view(studying[[1]])
studying_df_t1 <- as.data.frame(studying[[1]][1])
class(studying_df_t1)
# view(studying_df_t1)
weak_netw_t1 <- data.matrix(studying_df_t1)
weak_netw_t1[weak_netw_t1==9]<-NA
class(weak_netw_t1)
# view(weak_netw_t1)

studying_df_t2 <- as.data.frame(studying[[1]][2])
class(studying_df_t2)
# view(studying_df_t2)
weak_netw_t2 <- data.matrix(studying_df_t2)
weak_netw_t2[weak_netw_t2==9]<-NA
class(weak_netw_t2)
# view(weak_netw_t2)
```

### Network matrix (strong ties)
```{r}
# view(hanging[[1]])
hanging_df_t1 <- as.data.frame(hanging[[1]][1])
class(hanging_df_t1)
# view(hanging_df_t1)
strong_netw_t1 <- data.matrix(hanging_df_t1)
strong_netw_t1[strong_netw_t1==9]<-NA
class(strong_netw_t1)
# view(strong_netw_t1)

hanging_df_t2 <- as.data.frame(hanging[[1]][2])
class(hanging_df_t2)
# view(hanging_df_t2)
strong_netw_t2 <- data.matrix(hanging_df_t2)
strong_netw_t2[strong_netw_t2==9]<-NA
class(strong_netw_t2)
# view(strong_netw_t2)
```

### Covariates (gender, ethnicity)
```{r}
gender_df <- as.data.frame(gender)
ethnicity_df <- as.data.frame(ethnic)
ind_mat <- data.matrix(cbind(gender_df, ethnicity_df))
class(ind_mat)
# view(ind_mat)
```
## Visualisation
```{r}
require(igraph)

copy_strong_netw_t2 <- strong_netw_t2
copy_strong_netw_t2[is.na(copy_strong_netw_t2)] <- 0 

g_strong_t2 <- igraph:: graph_from_adjacency_matrix(copy_strong_netw_t2, mode = "directed")
plot(g_strong_t2)
```

## Explanation of NA's with fake data
```{r}
test_m <- matrix(NA, nrow = 4, ncol = 4)
test_m[1,1] <- 0
test_m[1,2] <- 1
test_m[1,4] <- 1
test_m[3,1] <- 0
test_m[3,2] <- 1
test_m[3,4] <- 1
test_m[4,1] <- 0
test_m[4,2] <- 1
test_m[4,4] <- 0

view(test_m)
```

For plotting what you do with missing data differs from how you deal with NA's than for RSiena

If someone isnt there in both the row and the column, you want them gone from your data
```{r}
misrow <- rowSums(is.na(test_m)) == nrow(test_m)
miscol <- colSums(is.na(test_m)) == nrow(test_m)

miss <- which(misrow & miscol) #Checken waar zowel de collumn als de rows missings zijn
miss

which(misrow) #Tells you which row has missing values

test_m[which(misrow), ] <- 0
```

<br>

# Webscraping
```{}
install.packages("xml2")
require(xml2)
install.packages("rvest")
require(rvest)
```
Trying to save data from a nos article
Saving the href

```{}
url <- "https://nos.nl/zoeken?q=Jochem+Tolsma&page=1"

nos <- read_html(url)
nos # It is a html document

lijst <- nos %>% html_elements("main") %>%
  html_elements("ul") %>%
  html_elements("li")

lijst[1] #shows that you can select just one of the loaded lists
# Keep digging deeper into these lists, until you reach the desired depth/ information (in this case the date)

# The href should be in the object under lijst 1
lijst[1] %>% html_children %>% html_attr(name = "href")

#Copy x-path, to lead R to the right place, on the right page
# //*[@id="content"]/div[2]

#Copy full x-path
# /html/body/div/main/div[2]
```
Saving the date
```{}
# By myself
time_num <- lijst[1] %>% html_children %>% html_children %>% .[2] %>% html_children %>% .[2] %>% html_children %>% html_children %>% html_attr(name = "datetime")
time_num

#.[2] --> selects the second list in the objects that are under the object that are under lijst 1

time_char <- lijst[1] %>% html_children %>% html_children %>% .[2] %>% html_children %>% .[2] %>% html_children %>% html_children %>% html_text 
time_char

# In class
lijst[1] %>% html_element("time") %>% html_attr(name = "datetime")
lijst[1] %>% html_element("time") %>% html_text
```

<br>

# Trying on my own (again)
## Descriptive statistics
```{r}
vcount(g_strong_t2)
ecount(g_strong_t2)

pretty_graph <- simplify(g_strong_t2)
pretty_graph <- g_strong_t2
plot(pretty_graph)

plot(pretty_graph,
     vertex.color = "pink", # nice color for the vertices
     vertex.size = 10,  # we'll vertices a bit smaller
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.family = "Helvetica", # not a fan of times new roman in figures
     vertex.label.cex = 0.3,  # make the label a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.075, # make arrow size (direction of edge) smaller
     margin = 0.01,
     rescale = FALSE, ylim = c(-9,9))
```
<br>

Giving gender different colours
```{r}
gender1 <- gender[[1]][,1]
gender2 <- gender[[1]][,2]
gender_test <- coalesce(gender1, gender2)
head(gender_test)
```
```{r}
hangingw2_nona <- na_if(hanging[[1]][[2]], 9)
#checking if one misses in rows
rowhangingw2 <- rowSums(is.na(hangingw2_nona)) == nrow(hangingw2_nona)
#checking if one misses in cols
colhangingw2 <- colSums(is.na(hangingw2_nona)) == ncol(hangingw2_nona)
hangingw2_nona[which(rowhangingw2) , ] <- 0

igraph::betweenness(pretty_graph, directed = TRUE)
V(pretty_graph)$size = betweenness(pretty_graph, normalized = T, directed = TRUE) * 100 + 10
pretty_graph_t2 <- igraph::graph_from_adjacency_matrix(hangingw2_nona, mode = "directed", weighted = NULL, diag = TRUE, add.colnames = NA,
    add.rownames = NA)

kleurtje <- ifelse(gender_test==2, "orange", "purple")
plot(pretty_graph_t2,
     vertex.color = kleurtje, # nice color for the vertices
     vertex.size = 5,  # we'll vertices a bit smaller
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.5,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)

pretty_graph_t2
```

Trying to change node size based on degree
```{r}
V(pretty_graph)$size = betweenness(pretty_graph, normalized = T, directed = T) * 1000 + 50

plot(pretty_graph,
     vertex.color = "pink",
     vertex.frame.color = "gray",
     vertex.label.color = "black",
     vertex.label.family = "Helvetica",
     vertex.label.cex = 0.3,
     vertex.label.dist = 0.5, 
     edge.curved = 0.2,
     edge.arrow.size = 0.075,
     margin = 0.01,
     rescale = FALSE, ylim = c(-9,9))

```

```{r}
V(pretty_graph)$color = betweenness(pretty_graph, normalized = T, directed = T) * 1000 + 50

plot(pretty_graph,
     vertex.size = 100,
     vertex.frame.color = "gray",
     vertex.label.color = "black",
     vertex.label.family = "Helvetica",
     vertex.label.cex = 0.3,
     vertex.label.dist = 0.5, 
     edge.curved = 0.2,
     edge.arrow.size = 0.075,
     margin = 0.01,
     rescale = FALSE, ylim = c(-9,9))
```

