---
title: "Appendix A SNASS"
author: "Lauren Dirix"
bibliography: references.bib
---

# Tutorial SNASS book
## A1 through A6 - Setting up
```{r}
rm(list = ls())

#install.packages("installr")
require(installr)
#install.packages("foreign")
require(foreign)
require(tidyverse)
#updateR()
```

Loading data sets
``` {r, warning = FALSE}
cv08 <- foreign::read.spss("Cultural_Changes_2008.sav", use.value.labels = T, to.data.frame = T)
cv10 <- foreign::read.spss("Cultural_Changes_2010.sav", use.value.labels = T, to.data.frame = T)

cv08_nolab <- foreign::read.spss("Cultural_Changes_2008.sav", use.value.labels = F, to.data.frame = T)
cv10_nolab <- foreign::read.spss("Cultural_Changes_2010.sav", use.value.labels = F, to.data.frame = T)

cv08_haven <- haven::read_spss("Cultural_Changes_2008.sav")
cv10_haven <- haven::read_spss("Cultural_Changes_2010.sav")
```

Inspect the data
```{r, include = FALSE}
str(cv08)
str(cv08_nolab)
str(cv08_haven)
```

```{r}
str(cv08$lftop)
str(cv08_haven$lftop)

attributes(cv08$lftop)
attributes(cv08_nolab$lftop)
attributes(cv08_haven$lftop)

names(cv08_haven)
summary(cv08_haven)
head(cv08_haven)
fix(cv08)
View(cv08_haven)
```

## A7 - Assign missing values
### In base
```{r}
str(cv08_haven$lftop)
summary(cv08_haven$lftop)
attr(cv08_haven$lftop, "labels")
table(cv08_haven$lftop, useNA = "always")

cv08$lftop_new <- cv08$lftop
cv08$lftop_new[cv08$lftop_new == "Onbekend"] <- NA

table(cv08$lftop_new, useNA = "always")
levels(cv08$lftop_new)

cv08$agen <- as.numeric(as.character(cv08$lftop_new))
table(cv08$agen, useNA = "always")

str(cv08$agen)
```

### In tidy
```{r}
#cv08_haven <- mutate(cv08_haven, lftop_new = lftop)
#cv08_haven$lftop_new <- na_if(cv08_haven$lftop_new, 99)

#OR (more quickly)

cv08_haven <- mutate(cv08_haven, lftop_new = na_if(lftop, 99))

table(cv08_haven$lftop_new, useNA = "always")
```


## A8 - Recoding variables 
### In base
```{r}
levels(cv08$var006n)
table(cv08$var006n, useNA = "always")

cv08$educn <- as.numeric(cv08$var006n)
table(cv08$educn, useNA = "always")

cv08$educ3 <- NA

cv08$educ3[cv08$educn == 2 | cv08$educn ==3] <- 1
cv08$educ3[cv08$educn > 3 & cv08$educn < 8] <- 2
cv08$educ3[cv08$educn > 7 & cv08$educn < 11] <- 3

table(cv08$educ3, useNA = "always")
prop.table(table(cv08$educ3, useNA = "always"))

cv08$educ3 <- as.factor(cv08$educ3)
table(cv08$educ3, useNA = "always")

levels(cv08$educ3) <- c( "primary", "secondary", "tertiary")
table(cv08$educ3, useNA = "always")
```
### In tidy
```{r}
#install.packages("labelled")
require(labelled)

str(cv08_haven$var006n)
attr(cv08_haven$var006n, "labels")

table(cv08_haven$var006n, useNA = "always")

cv08_haven <- mutate(cv08_haven, educ3 = recode(var006n, '-3' = -9, '-1' = 1, '1' = 1, '2' = 2, '3' = 2, '4' = 2, '5' = 2, '6' = 3, '7' = 3, '8' = 3, '10' = -9), .keep_value_labels = FALSE)
cv08_haven <- mutate(cv08_haven, educ3 = na_if(educ3, -9))

cv08_haven <- mutate(cv08_haven, educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))

table(cv08_haven$educ3, useNA = "always")
```
Using piping
```{r}
cv08_haven <- cv08_haven %>%
  mutate(cv08_haven, educ3 = recode(var006n, '-3' = -9, '-1' = 1, '1' = 1, '2' = 2, '3' = 2, '4' = 2, '5' = 2, '6'
  = 3, '7' = 3, '8' = 3, '10' = -9), .keep_value_labels = FALSE) %>%
  mutate(educ3 = na_if(educ3, -9)) %>%
  mutate(educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))
```

## A9 - Means and counting
### In base
```{r}
summary(cv08$int055)
summary(cv08$int056)
summary(cv08$int057)

cv08$int055n <- as.numeric(cv08$int055)
table(cv08$int055n, useNA = "always")
cv08$int056n <- as.numeric(cv08$int056)
table(cv08$int056n, useNA = "always")
cv08$int057n <- as.numeric(cv08$int057)
table(cv08$int057n, useNA = "always")

cv08$int055n[cv08$int055n < 5] <- NA
cv08$int055n <- cv08$int055n - 4
cv08$int056n[cv08$int056n < 5] <- NA
cv08$int056n <- cv08$int056n - 4
cv08$int057n[cv08$int057n < 5] <- NA
cv08$int057n <- cv08$int057n - 4

#means per respondent
testmeans <- rowMeans(cbind(cv08$int055n, cv08$int056n, cv08$int057n), na.rm = T)
head(testmeans)

nmis <- rowSums(is.na(cbind(cv08$int055n, cv08$ubt056n, cv08$int057n)))
testmeans <- ifelse(nmis <2, testmeans, NA)
cv08$int_mean <- testmeans
```

### In tidy
```{r}
table(cv08_haven$int055, useNA = "always")
table(cv08_haven$int056, useNA = "always")
table(cv08_haven$int057, useNA = "always")

cv08_haven <- cv08_haven %>%
  mutate(int055n = recode(int055, '-6' = -9, '-5' = -9, '-3' = -9, '-2' = -9, '1' = 4, '2' = 3, '3'
    = 2, '4' = 1), .keep_value_labels = FALSE) %>%
  mutate(int055n = na_if(int055n, -9)) %>%
  mutate(int055n = labelled(int055n, c('Helemaal geen tegenstelling' = 1,'Niet zo groot' = 2, 'groot'
    = 3, 'Zeer groot' = 4)))
cv08_haven <- cv08_haven %>%
  mutate(int056n = recode(int056, '-6' = -9, '-5' = -9, '-3' = -9, '-2' = -9, '1' = 4, '2' = 3, '3'
    = 2, '4' = 1), .keep_value_labels = FALSE) %>%
  mutate(int056n = na_if(int056n, -9)) %>%
  mutate(int056n = labelled(int056n, c('Helemaal geen tegenstelling' = 1,'Niet zo groot' = 2, 'groot'
    = 3, 'Zeer groot' = 4)))
cv08_haven <- cv08_haven %>%
  mutate(int057n = recode(int057, '-6' = -9, '-5' = -9, '-3' = -9, '-2' = -9, '1' = 4, '2' = 3, '3'
    = 2, '4' = 1), .keep_value_labels = FALSE) %>%
  mutate(int057n = na_if(int057n, -9)) %>%
  mutate(int057n = labelled(int057n, c('Helemaal geen tegenstelling' = 1,'Niet zo groot' = 2, 'groot'
    = 3, 'Zeer groot' = 4)))

  #mean regardless of how many missings
cv08_haven <- cv08_haven %>%
  rowwise() %>%
  mutate(int_mean = mean(c(int055n, int056n, int057n), na.rm = TRUE))

  #mean with max 1 missing, otherwise NA
cv08_haven <- cv08_haven %>%
  mutate(int_mean_temp = rowMeans(cbind(int055n, int056n, int057n), na.rm = TRUE), nmis = 
    rowSums(is.na(cbind(int055n, int056n, int057n))), int_mean = ifelse(nmis < 2, int_mean_temp, 
    NA)) %>% 
  dplyr:: select(-int_mean_temp, -nmis)
```

## A10 - Merging data files
### In base
1. Selecting the variables we want to keep
```{r}
cv08_sel <- cv08[, c("we_id", "lftop", "geslacht", "var006n", "v401", "landd")]
cv10_sel <- cv10[, c("Sleutel", "var002", "var001", "Vltoplop", "V401", "Landd")]
```

2. Make variables consistent
```{r}
names(cv08_sel) <- names(cv10_sel) <- c("id", "age", "sex", "educ", "health", "region")
summary(cv08_sel)
summary(cv10_sel)

#ID
cv08_sel$id <- as.numeric(as.character(cv08_sel$id))
cv10_sel$id <- as.numeric(as.character(cv10_sel$id))

#Age
cv08_sel$age <- as.numeric(as.character(cv08_sel$age))
cv10_sel$age <- as.numeric(as.character(cv10_sel$age))

#Sex
table(cv08_sel$sex, useNA = "always")
table(cv10_sel$sex, useNA = "always")

cv08_sel$sexn <- as.numeric(cv08_sel$sex)
table(cv08_sel$sexn, useNA = "always")
cv08_sel$men <- ifelse(cv08_sel$sexn == 2, 1, 0)
cv08_sel$men <- ifelse(cv08_sel$sexn == 1, NA, cv08_sel$men)
table(cv08_sel$men, useNA = "always")

cv10_sel$sexn <- as.numeric(cv10_sel$sex)
table(cv10_sel$sexn, useNA = "always")
cv10_sel$men <- ifelse(cv10_sel$sexn == 2, 1, 0)
table(cv10_sel$men, useNA = "always")

#Education
levels(cv08_sel$educ)
levels(cv10_sel$educ)
table(cv08_sel$educ, useNA= "always")
table(cv10_sel$educ, useNA= "always")

cv08_sel$educn <- as.numeric(cv08_sel$educ)
table(cv08_sel$educn)

cv08_sel$educ3 <- NA
cv08_sel$educ3[cv08_sel$educn == 2 | cv08_sel$educn == 3] <- 1
cv08_sel$educ3[cv08_sel$educn > 3 & cv08_sel$educn < 8] <- 2
cv08_sel$educ3[cv08_sel$educn > 7 & cv08_sel$educn < 11] <- 3
table(cv08_sel$educ3, useNA = "always")

cv10_sel$educn <- as.numeric(cv10_sel$educ)
table(cv10_sel$educn)

cv10_sel$educ3 <- NA
cv10_sel$educ3[cv10_sel$educn < 3] <- 1
cv10_sel$educ3[cv10_sel$educn > 2 & cv10_sel$educn < 6] <- 2
cv10_sel$educ3[cv10_sel$educn == 6] <- 3
table(cv10_sel$educ3, useNA = "always")
```
3. Merge
```{r}
cv08_sel$wave <- 2008
cv10_sel$wave <- 2010

cv08_sel$id2 <- rank(cv08_sel$id)
cv10_sel$id2 <- rank(cv10_sel$id)

#clean up data sets
cv08_sel <- cv08_sel[, c("id", "id2", "age", "men", "educ3", "health", "region", "wave")]
cv10_sel <- cv10_sel[, c("id", "id2", "age", "men", "educ3", "health", "region", "wave")]

summary(cv08_sel)
summary(cv10_sel)

#Long format
cv_tot <- rbind(cv08_sel, cv10_sel)
summary(cv_tot)

#Wide format
cv_tot_panel <- merge(cv08_sel, cv10_sel, all = TRUE, by = "id2")
head(cv_tot_panel)
```

### In tidy
```{r}
cv08_sel <- cv08_haven %>%
  dplyr:: select(c("we_id", "lftop", "geslacht", "var006n", "v401", "landd"))
cv10_sel <- cv10_haven %>%
  dplyr:: select(c("Sleutel", "var002", "var001", "Vltoplop", "V401", "Landd"))

names(cv08_sel) <- names(cv10_sel) <- c("id", "age", "sex", "educ", "health", "region")

cv08_sel <- cv08_sel %>%
    mutate(age = na_if(age, 99))
cv10_sel <- cv10_sel %>%
    mutate(age = na_if(age, 99))

cv08_sel <- cv08_sel %>%
    mutate(men = recode(sex, `9` = -9, M = 1, V = 0, .keep_value_labels = FALSE), 
      men = na_if(men, -9), men = labelled(men, c(man = 1, vrouw = 0)))
cv10_sel <- cv10_sel %>%
    mutate(men = recode(sex, `2` = 0, .keep_value_labels = FALSE), men = labelled(men, 
      c(man = 1, vrouw = 0)))

cv08_sel <- cv08_sel %>%
    mutate(educ3 = recode(educ, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2, `4` = 2, `5` = 2, 
      `6` = 3, `7` = 3, `8` = 3, `10` = -9, .keep_value_labels = FALSE), educ3 = na_if(educ3, -9), 
      educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))

cv10_sel <- cv10_sel %>%
    mutate(educ3 = recode(educ, `-1` = 1, `1` = 1, `2` = 2, `3` = 2, `4` = 2, `5` = 3, `10` = -9, 
      .keep_value_labels = FALSE), educ3 = na_if(educ3, -9), educ3 = factor(educ3, levels = c(1, 2,
      3), labels = c("primary", "secondary", "tertiary")))

cv08_sel$wave <- 2008
cv10_sel$wave <- 2010

cv08_sel$id2 <- rank(cv08_sel$id)
cv10_sel$id2 <- rank(cv10_sel$id)

cv08_sel <- cv08_sel %>%
    dplyr::select(c("id", "id2", "age", "men", "educ3", "health", "region", "wave"))

cv10_sel <- cv10_sel %>%
    dplyr::select(c("id", "id2", "age", "men", "educ3", "health", "region", "wave"))

#Merge long
cv_tot_tidy <- cv08_sel %>%
    add_row(cv10_sel)

summary(cv_tot_tidy)

#Merge wide
cv_tot_panel_tidy <- full_join(cv08_sel, cv10_sel, by = "id2", suffix = c(".2008", ".2010"))
```

## A11 - Aggregate data
### In base
```{r}
age_region <- aggregate(cv_tot$age, by = list(cv_tot$region), FUN = mean)
head(age_region)

age_region <- aggregate(cv_tot$age, by = list(cv_tot$region), FUN = mean, na.rm = TRUE)
head(age_region)
names(age_region) <- c("region", "age_region")
age_region

cv_total <- merge(cv_tot, age_region, by = "region", all.x = TRUE)
```

### In tidy
```{r}
age_region <- group_by(cv_tot_tidy, region)
age_region <- mutate(age_region, mean = mean(age, na.rm = TRUE))

# Or shorter:
cv_total_tidy <- cv_tot_tidy %>%
    group_by(region) %>%
    mutate(age_region = mean(age, na.rm = TRUE)) %>%
    ungroup()
```

## A12 - Missing values
### In base
```{r}
cv_total$health[cv_total$health == "Weigert"] <- NA
cv_total$healthn <- as.numeric(cv_total$health) - 4
table(cv_total$health, useNA = "always")
table(cv_total$healthn, useNA = "always")
```
Listwise deletion
```{r}
summary(cv_total)
model2 <- lm(as.numeric(healthn) ~ men + age + as.factor(educ3) + as.factor(region), data = cv_total)
summary(model2)
```
Replacing missing values
```{r}
# Categorical variables (add NA category)
cv_total$men2 <- ifelse(is.na(cv_total$men), 2, cv_total$men)
summary(cv_total$men2)

cv_total$educ3b <- ifelse(is.na(cv_total$educ3), 4, cv_total$educ3)
summary(cv_total$educ3b)

# Continuous variables (input the mean)
cv_total$age2 <- ifelse(is.na(cv_total$age), mean(cv_total$age, na.rm = TRUE), cv_total$age)
cv_total$age_mis <- ifelse(is.na(cv_total$age), 1, 0)
summary(cv_total$age2)
table(cv_total$age_mis)

model3 <- lm(healthn ~ as.factor(men2) + age2 + age_mis + as.factor(educ3b) + as.factor(region), data = cv_total)
summary(model3)
```
Imputing missing values
```{r}
# Reset data set
cv_total <- cv_total[, c("id", "id2", "age", "men", "educ3", "health", "region", "wave")]
cv_total$health[cv_total$health == "Weigert"] <- NA
cv_total$health <- as.numeric(cv_total$health) - 4

cv_total$men <- as.factor(cv_total$men)
cv_total$educ3 <- as.factor(cv_total$educ3)

# install.packages("mice")
require(mice)
mice:: md.pattern(cv_total)

imp <- mice(data = cv_total, method = c("", "", "cart", "logreg", "polr", "cart", "", ""))
attributes(imp)
pred <- imp$pred
pred[, "id"] <- 0
pred[, "id2"] <- 0
pred[, "wave"] <- 0
pred

imp <- mice(data = cv_total, method = c("", "", "cart", "logreg", "polr", "cart", "", ""), pred = pred, seed = 45622)

summary(cv_total)
summary(complete(imp))

plot(imp)

model_imp <- with(imp, lm(as.numeric(health) ~ men + age + educ3 + region))
pool_model_imp <- pool(model_imp)
summary(pool_model_imp)
```
### In tidy
```{r}
cv_total_tidy <- cv_total_tidy %>%
    mutate(health = na_if(health, -2), healthn = as.numeric(health) - 4)
```

Listwise deletion
```{r}
model2 <- cv_total_tidy %>%
    mutate(across(c(educ3, region), as.factor)) %>%
    lm(formula = healthn ~ men + age + educ3 + region) %>%
    summary()

model2
```
Replacing missing values
```{r}
cv_total_tidy2 <- cv_total_tidy %>%
    mutate(across(everything(), as.numeric), men2 = replace_na(men, 2), educ3b = replace_na(educ3,
    4), age2 = replace_na(age, mean(age, na.rm = TRUE)), age_mis = ifelse(is.na(age), 1, 0))

model3 <- cv_total_tidy2 %>%
    mutate_at(c("men2", "educ3b", "region"), as.factor) %>%
    lm(formula = healthn ~ men2 + age2 + age_mis + educ3b + region) %>%
    summary()

model3
```
Imputing missing values
```{r}
pred <- cv_total_tidy %>%
  dplyr::select(c("id", "id2", "age", "men", "educ3", "health", "region", "wave")) %>%
  mutate_all(as.numeric) %>%
  mutate_at(c("men", "educ3"), as.factor) %>%
  mice(method = c("", "", "cart", "logreg", "polr", "cart", "", "")) %>%
  unclass() %>% 
  pluck("predictorMatrix") %>%
  as.data.frame() %>% 
  mutate_at(c("id", "id2", "age"), funs(recode(1, 0))) %>% 
  as.matrix()

model_imp <- cv_total_tidy %>%
  dplyr::select(c("id", "id2", "age", "men", "educ3", "health", "region", "wave")) %>%
  mutate_all(as.numeric) %>%
  mutate_at(c("men", "educ3"), as.factor) %>%
  mice(method = c("", "", "cart", "logreg", "polr", "cart", "", ""), pred = pred, seed = 45622) %>%
  with(lm(as.numeric(health) ~ men + age + educ3 + region)) %>%
  pool()

summary(model_imp)
```