---
title: "Week 3"
author: "Lauren Dirix"
bibliography: references.bib
---

# Preparation before class

## Adjusting my research questions

My descriptive, macro, question was: "Does positive peer influence affect youths' political participation?" After our discussion in class, I realise that this isn't really a descriptive question, as it is about a relationship between two variables. So I will adjust my research question to be *actually* descriptive, at least I hope so. Also, with an eye on the available data, I have decided to slightly adjust my topic. Rather than looking at political participation, I will look at civic engagement. My adjusted research question is: *Are students who are connected to each other more similar with regards to civic engagement, than students who are not connected to each other?*

It may be interesting here to look at both strong and weak ties, as in the data there is a measure for 'friends' and 'people you study with.' However, I don't know whether that will be too complicated for this course yet.

My explanatory, micro, question was: "Do peers become more similar to each other with regards to the degree of political participation?" I think I will keep this question relatively similar to what it was, except, of course, adjusting the topic to fit with my new macro question. I am still interested in seeing whether potential similarity can be explained through selection effects or through influence effects. So my 'new' research question is: *Do students become more similar to each other with regards to civic engagement, when they are connected in a network?*

Here again, it could be really interesting to look at the difference between strong and weak ties, as you could expect there to be a bigger effect of strong ties, if there is an influence effect.

<br>

## What data do I need?

I think I will use the MRCS data, as this is a very extensive dataset, which covers the topic that I am interested in.

**Civic engagement**

For civic engagement they asked questions about both beliefs and behaviour. About beliefs they asked:

- SXceng01: I feel responsible for my local community.

- SXceng02: I believe I should make a difference in my local community.

- SXceng03: I am committed to serve in my local community.

- SXceng04: I know how to give back to my local community

About behaviour they asked:

- SXceng05: I am involved in structured volunteer position(s) in my local community.

- SXceng06: When working with others, I make positive changes in my local community.

- SXceng07: I help members of my local community.

- SXceng08: I stay informed of events in my local community.

- SXceng09: I participate in discussions that raise issues of social responsibility.

- SXceng10: I contribute to charitable organizations within the local community

I am most interested in the actual behaviour, however, I don't know whether this behaviour is salient enough in the data to see any patterns.

**Strong/ weak ties** 

To map the social network within the learning communities, the researchers asked multiple questions. Firstly, for what I will consider weak ties, the respondents were asked to write down who they tend to study with.

Who do you study with in your MLC? Please write their names in the box below. If you do not know their
last names, please include as much as you can remember. (If you can’t think of anyone, please skip this question)

- SXncst01-10: Write in the first and last name

And for the measure of strong ties, I will be relying on the following question:

Which students in your MLC do you hang around with and talk to the most? Please write their names
in the box below. Again, if you do not know their last names, please include as much as you can
remember. (If you can't think of anyone, please skip this question)

- SXnhng01-10: Write in the first and last name

------------------------------------------------------------------------

# RSiena workshop
## Logic
Set up
```{r}
rm(list = ls())

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

#Saves the data automatically with a date stamp

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}
```

```{r}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
# devtools::install_github('JochemTolsma/RsienaTwoStep', build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```

```{r}
ts_net1
net1g <- graph_from_adjacency_matrix(ts_net1, mode = "directed")
coords <- layout_(net1g, nicely())  #let us keep the layout
par(mar = c(0.1, 0.1, 0.1, 0.1))
{
    plot.igraph(net1g, layout = coords)
    graphics::box()
}
```

Ego-sampling
```{r}
set.seed(24553253)
ego <- ts_select(net = ts_net1)
ego
```
```{r}
options <- ts_alternatives_ministep(net = ts_net1, ego = ego)
# options
plots <- lapply(options, graph_from_adjacency_matrix, mode = "directed")
par(mar = c(0, 0, 0, 0) + 0.1)
par(mfrow = c(2, 2))

fplot <- function(x) {
    plot.igraph(x, layout = coords, margin = 0)
    graphics::box()
}

lapply(plots, fplot)
```

Network statistics
```{r}
ts_degree(net = options[[1]], ego = ego)
sapply(options, ts_degree, ego = ego)
sapply(options, ts_recip, ego = ego)
```
First line: Shows how many outdegrees actor 4 (= ego) has in the first simulated network (after ministep 1)
Second line: Shows how many outdegrees actor 4 (= ego) has in all simulated networks (after ministep 1)
Third line: Shows how many reciprocated ties actor 4 (= ego) has in all simulated networks (after ministep 1)

```{r}
option <- 4
ts_degree(options[[option]], ego = ego) * -1 + ts_recip(options[[option]], ego = ego) * 1.5
```
For option 4, actor 4 has 1 tie, which get a value (parameter) of -1, but he also gets one reciprocated tie, which we value 1.5 --> actor 4 in option 4 gets a value of .5.
This is just to show, you would not fill out the values for -1 and 1.5. This is what the model gives you.

```{r}
eval <- sapply(options, FUN = ts_eval, ego = ego, statistics = list(ts_degree, ts_recip), parameters = c(-1,
    1.5))
eval
print("network with maximum evaluation score:")
which.max(eval)
```

This prints all options, in which the parameters are taken into account. It shows which one has the best value for this specific actor. It shows the preference for each network.
To determine which network the agent actually chooses, you need to weigh them --> you need to recalculate all values to end up between 0 and 1, all of which chances have to sum to 1
For every network, a chance of choosing this network is calculated using McFaddens choice function

```{r}
choice <- sample(1:length(eval), size = 1, prob = exp(eval)/sum(exp(eval)))
print("choice:")
choice
# print('network:') options[[choice]]
```

Because every choice is dependent on the previous choices made, there is interdependency in the network

## Analyses
Set up
```{r}
df <- fload("./data/20230621df_complete.rda")
publications <- fload("./data/20230621list_publications_jt.rda")

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}
```

```{r}
packages = c("RSiena", "tidyverse", "stringdist", "stringi")

fpackage.check(packages)
```
```{r}
output <- f_pubnets()
df_soc <- output[[1]]
df_network <- output[[2]]
```

Start here for working with RSiena
### Step 1
df_soc --> dataset on actor level
df_network --> dataset on network level (the dependent variable/ Y)

```{r}
view(df_soc)
view(df_network)
```

The data has multiple waves, seperate those

```{r}
wave1 <- df_network[[1]]
wave2 <- df_network[[2]]

# let us put the diagonal to zero
diag(wave1) <- 0
diag(wave2) <- 0

# we want a binary tie (not a weighted tie)
wave1[wave1 > 1] <- 1
wave2[wave2 > 1] <- 1

# put the nets in an array
net_soc_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

# dependent
net <- sienaDependent(net_soc_array)
```

RSiena can only handle numeric data, so recode gender (= the independent variable)

```{r}
# gender (= independent variable)
gender <- as.numeric(df_soc$gender == "female")
gender <- coCovar(gender)
```

```{r}
mydata <- sienaDataCreate(net, gender)
```

### Step 2
```{r}
myeff <- getEffects(mydata)
# effectsDocumentation(myeff)
```
Effects documentation tells you what network statistics you can calculate

### Step 3
Get a quick overview of the dataframe (automatically saves it into a document into a new map of the journal)

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

print01Report(mydata, modelname = "./results/soc_init")
```

### Step 4
You want to safe the effects that you want to consider (based on theory)
Hint:make every effect a seperate line, so you can easily change it

```{r}
myeff <- includeEffects(myeff, isolateNet, inPop, outAct, inAct, transTrip)
```

```{r}
myeff <- includeEffects(myeff, sameX, egoX, altX, interaction1 = "gender")
```
Interaction1 means that you estimate for the variable afterwards (in this case gender)

### Step 5
You can decide to use different algorithms, or multiple cores or whatever, but here just the basic
```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "soc_init")
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff))
# (the outer parentheses lead to printing the obtained result on the screen) if necessary, estimate
# further
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans, returnDeps = TRUE))
```
What does this mean:
Rate parameter: On average every agent has made 4 ministeps
Outdegree (density): Very strongly negative
Network isolate effect: There are probably a lot of people who do not have any ties --> you get a positive score if there are neither indegrees nor outdegrees
Eval gender alter: Men send fewer ties
Eval gender ego: Women send more ties
Eval same gender: Positive, so people prefer to work together with people of the same gender

### Step 6
You want to know whether your simulated network matches your observed network. You make the simulated network match on certain criteria, but a network has more characteristics (e.g. degree distribution). This is interesting to see whether it matches as well. 
```{r}
# see here: ?'sienaGOF-auxiliary'

# The geodesic distribution is not available from within RSiena, and therefore is copied from the
# help page of sienaGOF-auxiliary:

# GeodesicDistribution calculates the distribution of non-directed geodesic distances; see
# ?sna::geodist The default for \code{levls} reflects the usual phenomenon that geodesic distances
# larger than 5 do not differ appreciably with respect to interpretation.  Note that the levels of
# the result are named; these names are used in the \code{plot} method.
GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

# The following function is taken from the help page for sienaTest

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

```{r}
gofi0 <- sienaGOF(ans, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0 <- sienaGOF(ans, OutdegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gof0.gd <- sienaGOF(ans, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
gof0.tc <- sienaGOF(ans, TriadCensus, verbose = FALSE, join = TRUE, varName = "net")
```

```{r}
plot(gofi0)
```
This plot shows the observerd value (in red) and the distribution of simulated networks (the boxplots). On the x-axis it shows the amount of agents in a network with X amount of ties
Example: For 0 ties, the model estimates fewer people having 0 ties, than actually observed
P value: Measures whether overall the deviation is significant (in this case, not significant)

### Step 7
Can be used to figure out how often a certain statistic/ effect has played a role in a decision from an agent
Would an agent have made a different choice, if the statistic would have been 0
Shows how important a specific statistic is, relative to other ties
```{r}

```

### Step 8
Using RSiena as an ABM/ making the micro macro link
```{r}

```

<br>

# On own data
Getting started
```{r}
rm(list = ls())

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}
```

Install/ load packages
```{r}
packages = c("RSiena", "tidyverse", "stringdist", "stringi")

fpackage.check(packages)
```


